Can Auxiliary Indicators Improve COVID-19 Forecasting and Hotspot Prediction?

### 2021-06-21 forecast/paper/paper.tex
Rebuilt figs; finished supplement draft; added Justin's suggestion

\begin{figure}[tb!]
  \includegraphics[width=\columnwidth]{fig/revisions.pdf}
  % RJT: Ensuring we start with a "declarative title" :) 
  % https://www.internationalscienceediting.com/how-to-write-a-figure-caption/
  \caption{Revision behavior for two signals in the HRR containing New York  
    City.  Each colored line corresponds to the data as reported on a particular 
    date (\textit{as of} dates varying from September 28 through October
    19). The left panel shows the DV-CLI signal, which was regularly revised
    throughout the period, though the effects fade as we look further back in   
    time. In contrast, the right panel shows case rates reported by JHU CSSE
    (smoothed with a 7-day trailing average), which remain ``as reported'' on
    September 28, with a spike towards the end of this period, until a major
    correction is made on October 19, which brings this down and affects all
    prior data as well.}  
  \label{fig:vintage}
\end{figure}

@@

One might ask if the benefits we observe in forecasting and hotspot
prediction have anything to do with the actual indicator themselves. A plausible 
alternative explanation is that the indicators are simply providing \textit{implicit
  regularization} on top of  the basic AR model when we include them as lagged
features in \eqref{eq:forecast_ar_x} and \eqref{eq:hotspot_ar_x}.

>>>>

\begin{figure}[tb!]
  \includegraphics[width=\columnwidth]{fig/revisions.pdf}
  % RJT: Ensuring we start with a "declarative title" :) 
  % https://www.internationalscienceediting.com/how-to-write-a-figure-caption/
  \caption{Revision behavior for two signals in the HRR containing Charlotte,
    North Carolina.  Each colored line corresponds to the data as reported on a
    particular 
    date (\textit{as of} dates varying from September 28 through October
    19). The left panel shows the DV-CLI signal, which was regularly revised
    throughout the period, though the effects fade as we look further back in   
    time. In contrast, the right panel shows case rates reported by JHU CSSE
    (smoothed with a 7-day trailing average), which remain ``as reported'' on
    September 28, with a spike towards the end of this period, until a major
    correction is made on October 19, which brings this down and affects all
    prior data as well.}  
  \label{fig:vintage}
\end{figure}

@@

One might ask if the benefits we observe in forecasting and hotspot
prediction have anything to do with the actual indicator themselves. A plausible  
alternative explanation is that the indicators are simply providing \textit{implicit
  regularization} on top of the basic AR model, in the same way any noise
variable might, when we include them as lagged
features in \eqref{eq:forecast_ar_x} and \eqref{eq:hotspot_ar_x}.

### 2021-06-21 forecast/paper/supplement-text.tex
Rebuilt figs; finished supplement draft; added Justin's suggestion

Currently, both figures are in the manuscript. Probably just need text here.

>>>>

In Section 2.D of the main text, we discuss the extent to which the indicators
are leading or lagging case rates during different periods. To define the amount
of leadingness or laggingness, we use the cross correlation function (CCF)
between the two time series. The CCF is of an indicator $X_{\ell,t}$ and case
rates $Y_{\ell,t}$ is defined as
\[
\CCF_{\ell,t}(a) = \frac{1}{n_a} \sum_{i=1}^{n_a} \left( ( X_{\ell,t,i+a} -
  \overline{X}_{\ell,t}) / s^{X}_{\ell,t}\right) \left( ( Y_{\ell,t,i} -
  \overline{Y}_{\ell,t}) / s^{Y}_{\ell,t}\right), 
\]
where $n_a$ is the number of available time points when $X_{\ell,t}$ has been
shifted in time by $a$ days, and $\overline{X}_{\ell,t}$, $s^X_{\ell,t}$
are the sample mean and
standard deviation of $X_{\ell,t}$ (respectively $Y_{\ell,t}$). The result is a
sequence of Pearson correlations for each selected $a$, where we align
$X_{\ell,t}$ with the values of $Y_{\ell,t}$ that occurred $a$ days earlier.
Thus, for $a>0$, $\CCF_{\ell,t}(a)>0$ indicates that
$Y_{\ell,t}$ is moving together with the values of $X_{\ell,t}$ that occur $a$
days in the future. In this case we say that $X_{\ell,t}$ is lagging
$Y_{\ell,t}$. For $a<0$, $\CCF_{\ell,t}(a)>0$ means that $Y_{\ell,t}$ is
positively correlated with $X_{\ell,t}$ shifted earlier, so we say that
$X_{\ell,t}$ leads $Y_{\ell,t}$.

Figure \ref{fig:ccf-dv-finalized} shows the standardized signals for the HRR
containing Charlotte, North Carolina, from August 1, 2020 until the end of
September. These are the same signals shown in Figure 1 in the manuscript. To
define ``leadingness'' we compute $\CCF_{\ell,t}(a)$ for each $a\in
\{-15,\ldots,15\}$ using the 56 days leading up to time $t$ (this is the same
amount of data used to train the forecasters: 21 days of training data, 21 days
to get the response at $a=21$, and 14 days for the longest lagged value). The
orange dashed horizontal line represents the 97.5\% significance threshold for
correlations based on 56 observations. Any correlations larger in magnitude than
this value are considered statistically significant under the null hypothesis of
no relationship. We define leadingness to be the sum of the significant
correlations that are leading (those above the dashed line with $a<0$) while
laggingness is the same but for $a>0$. In this case, there are no significant
correlations, so both scores will be 0: on September 28, \dv~is neither
leading nor lagging cases.

Figure \ref{fig:lagging-only} shows the correlation between laggingness and the
difference in indicator WIS and AR WIS. Unlike leadingness (Figure 5 in the
manuscript) there is no obvious relationship. This is heartening as laggingness
should not aid forecasting performance. On the other hand, if an indicator is
more lagging than it is leading, this may suggest diminished performance.
Figure~\ref{fig:diff-in-lead-lag} shows the correlation of the difference in
leadingness and laggingness with the difference in WIS. The pattern here is
largely similar to the pattern in leadingness described in the manuscript: the
relationship is strongest in down periods and weakest in up periods with the
strength diminishing as we move from down to flat to up for all indicators.

### 2021-06-21 forecast/paper/supplement.Rmd
typos and minor supplement corrections
Rebuilt figs; finished supplement draft; added Justin's suggestion

```{r lagging-only, eval=FALSE}
# Deprecated
df2 %>%
  filter(period != "jm") %>%
  group_by(forecaster, udf) %>%
  summarise(
    leadingness = cor(AR_wis - wis, leading, use = "pairwise.complete.obs"),
    laggingness = cor(AR_wis - wis, lagging, use = "pairwise.complete.obs")) %>%
  ggplot(aes(udf, laggingness, fill = forecaster)) +
  geom_bar(width = 0.6, position = position_dodge(width=0.6),
           stat = "identity") +
  geom_hline(yintercept = 0) +
  scale_fill_manual(values = fcast_colors) +
  theme_bw() +
  ylab("Correlation") +
  xlab("") +
  theme(legend.position = "bottom", legend.title = element_blank())
```

```{r diff-in-lead-lag, include=FALSE}
# Figure 7 in the manuscript
df2 %>%
  filter(period != "jm") %>%
  group_by(forecaster, udf) %>%
  summarise(cor = cor(AR_wis - wis, leading - lagging, 
                      use = "pairwise.complete.obs")) %>%
  ggplot(aes(udf, cor, fill = forecaster)) +
  geom_bar(width = 0.6, position = position_dodge(width=0.6),
           stat = "identity") +
  geom_hline(yintercept = 0) +
  scale_fill_manual(values = fcast_colors) +
  theme_bw() +
  ylab("Correlation") +
  xlab("") +
  theme(legend.position = "bottom", legend.title = element_blank())
```

@@

```{r ccf-dv-finalized, eval=FALSE}
# Figure ?? in the manuscript
source("../code/ccf-dv-finalized.R")
gg
```

@@

\movie{Type legend for the movie here.}

>>>>

```{r ccf-dv-finalized, eval=TRUE, fig.cap="Illustration of the cross-correlation function between DV-CLI and cases. The left panel shows the standardized signals over the period from August 1 to September 28 (as of September 28, 2020). The right panel shows $\\CCF(a)$ for different values of $a$ as vertical blue bars. The orange dashed lines indicate the 95\\% significance threshold. By our leadingness metric, DV-CLI is neither leading nor lagging cases over this period."}
source("../code/ccf-dv-finalized.R")
gg
```

```{r lagging-only, fig.cap="Correlation of the difference in WIS with the  laggingness of the indicator at the target date, stratified by up, down, or flat period. Compare to Figure 5 in the manuscript."}
# Deprecated
df2 %>%
  filter(period != "jm") %>%
  group_by(forecaster, udf) %>%
  summarise(
    leadingness = cor(AR_wis - wis, leading, use = "pairwise.complete.obs"),
    laggingness = cor(AR_wis - wis, lagging, use = "pairwise.complete.obs")) %>%
  ggplot(aes(udf, laggingness, fill = forecaster)) +
  geom_bar(width = 0.6, position = position_dodge(width=0.6),
           stat = "identity") +
  geom_hline(yintercept = 0) +
  scale_fill_manual(values = fcast_colors) +
  theme_bw() +
  ylab("Correlation") +
  xlab("") +
  theme(legend.position = "bottom", legend.title = element_blank())
```

```{r diff-in-lead-lag, fig.cap = "Correlation of the difference between leadingness and laggingness with the difference in WIS. The relationship is essentially the same as described in the manuscript and shown in Figure 5."}
# Deprecated
df2 %>%
  filter(period != "jm") %>%
  group_by(forecaster, udf) %>%
  summarise(cor = cor(AR_wis - wis, leading - lagging, 
                      use = "pairwise.complete.obs")) %>%
  ggplot(aes(udf, cor, fill = forecaster)) +
  geom_bar(width = 0.6, position = position_dodge(width=0.6),
           stat = "identity") +
  geom_hline(yintercept = 0) +
  scale_fill_manual(values = fcast_colors) +
  theme_bw() +
  ylab("Correlation") +
  xlab("") +
  theme(legend.position = "bottom", legend.title = element_blank())
```

@@

@@

### 2021-06-21 forecast/paper/supplement.tex
typos and minor supplement corrections
Rebuilt figs; finished supplement draft; added Justin's suggestion

>>>>

{\centering \includegraphics[width=\textwidth]{fig/ccf-dv-finalized-1}

\caption{Illustration of the cross-correlation function between DV-CLI and cases. The left panel shows the standardized signals over the period from August 1 to September 28 (as of September 28, 2020). The right panel shows $\CCF(a)$ for different values of $a$ as vertical blue bars. The orange dashed lines indicate the 95\% significance threshold. By our leadingness metric, DV-CLI is neither leading nor lagging cases over this period.}\label{fig:ccf-dv-finalized}
\end{figure}

{\centering \includegraphics[width=\textwidth]{fig/lagging-only-1}

\caption{Correlation of the difference in WIS with the  laggingness of the indicator at the target date, stratified by up, down, or flat period. Compare to Figure 5 in the manuscript.}\label{fig:lagging-only}
\end{figure}

{\centering \includegraphics[width=\textwidth]{fig/diff-in-lead-lag-1}

\caption{Correlation of the difference between leadingness and laggingness with the difference in WIS. The relationship is essentially the same as described in the manuscript and shown in Figure 5.}\label{fig:diff-in-lead-lag}
\end{figure}
